---
title: "Practical Machine Learning"
subtitle: "Quiz1"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Question 1

Which of the following are components in building a machine learning algorithm?

Artificial intelligence

Statistical inference

Training and test sets

Deciding on an algorithm <-

Machine learning

## Question 2

Suppose we build a prediction algorithm on a data set and it is 100% accurate on that data set. Why might the algorithm not work well if we collect a new data set?

Our algorithm may be overfitting the training data, predicting both the signal and the noise. <-

We have used neural networks which has notoriously bad performance.

We may be using bad variables that don't explain the outcome.

We have too few predictors to get good out of sample accuracy.

## Question 3

What are typical sizes for the training and test sets?

100% training set, 0% test set.

90% training set, 10% test set

20% test set, 80% training set.

60% in the training set, 40% in the testing set. <--

## Question 4

What are some common error rates for predicting binary variables (i.e. variables with two possible values like yes/no, disease/normal, clicked/didn't click)? Check the correct answer(s).

Root mean squared error

R^2

Specificity <-

Correlation

Median absolute deviation

## Question 5

Suppose that we have created a machine learning algorithm that predicts whether a link will be clicked with 99% sensitivity and 99% specificity. The rate the link is clicked is 1/1000 of visits to a website. If we predict the link will be clicked on a specific visit, what is the probability it will actually be clicked?

99%

9% <-

89.9%

0.009%
```{r}
# 100,000 visits => 100 clicks
# 99% = sensitivity = TP/(TP+FN) = 
99/(99+1)
# 99% specificity =TN/(TN+FP) = 
98901/(98901+999)
# P(actually clicked|clicked) = TP/(TP+FP) = 
99/(99+999)
```
